Pretty darn good control: when are approximate solutions better than approximate models

### 2023-01-30 manuscript/manuscript.Rmd
hand-wavy footnote

One intuition for why a constant escapement policy performs so well in such a model comes from the anticorrelation between $X$ and $Y$ noted in the results section.
We have observed this pattern arise in a variety of policies, both classical as well as some obtained by training DRL agents.
In fact, ‘‘crashes’’ in the $X$ population are almost invariably shortly preceeded by dramatic increases in $Y$.
We hypothesize that this effectively leads to a dimensionality reduction: once the value of $X$ is known, along a path traced by the control system, then $Y$ may be roughly predicted.
A similar, although weaker, anticorrelation is also observed between $X$ and $Z$.
These two informal observations point at an effective simplification of the prediction process: the values of $Y$ and $Z$ are much less impactful to policy decisions than the value of $X$ itself.
We have not yet pushed further to establish a more precise picture of this dynamical simplification.

>>>>

One intuition for why a constant escapement policy performs so well in such a model comes from the anti-correlation between $X$ and $Y$ noted in the results section.
We have observed this pattern arise in a variety of policies, both classical as well as some obtained by training DRL agents.
In fact, ‘‘crashes’’ in the $X$ population are almost invariably shortly preceeded by dramatic increases in $Y$.
We hypothesize that this effectively leads to a dimensionality reduction: once the value of $X$ is known, along a path traced by the control system, then $Y$ may be roughly predicted.
A similar, although weaker, anti-correlation is also observed between $X$ and $Z$.
These two informal observations point at an effective simplification of the prediction process: the values of $Y$ and $Z$ are much less impactful to policy decisions than the value of $X$ itself.
We have not yet pushed further to establish a more precise picture of this dynamical simplification.\footnote{
A hand-waving way to understand this phenomenon is the following:
The $X$-$Y$ anti-correlation can be associated to the Lotka-Volterra competition term between $X$ and $Y$, i.e. $c_{XY}XY$ in eq. \eqref{eq:3d model stoch}.
Now, consider a fall in population $X$.
Because of the aforementioned anti-correlation with $Y$, this is accompanied by a growth in $Y$.
We may understand this, roughly, as a \emph{replacement} of $X$ for $Y$.
But then, because the parameter $D = 1.1 > 1$, this increases the effective birth rate $Z$, i.e. the term $f(X+DY)Z$ in eq. \eqref{eq:3d model stoch}.
This, in turn, leads to a growth of $Z$ and, thus, an anti-correlation between $Z$ and $X$.}
